---
output:
  bookdown::html_document2:
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r echo = FALSE, cache = FALSE}
source("utils.R", local = TRUE)
```

# Introduction to R {#CHAPTER-1}

R is a programming language and software environment for statistical computing
and graphics. It is widely used among statisticians and data scientists for data
analysis and visualization. However, R can also be used in an auditing context.
R has a large and active community of users, and there are many resources
available for learning and using R.

## Basics

One of the basic features of R is its ability to manipulate data. In R, basic
calculations work by using the standard arithmetic operators such as `+` for
addition, `-` for subtraction, `*` for multiplication, and `/` for division.
For example, if you want to calculate 2 + 3, you would type in `2 + 3` and R
will return the result of 5.

```{r}
2 + 3
```

R also allows for more advanced calculations such as exponentiation using the
`^` operator, and square roots using the `sqrt()` function. For example, to
calculate the square root of 9, you would type in `sqrt(9)` and R will return
the result of 3.

```{r}
sqrt(9)
```

You can also use parentheses to specify the order of operations in your
calculations. For example, if you want to calculate (2 + 3) * 4, you would type
in `(2 + 3) * 4` to get the result of 20.

```{r}
(2 + 3) * 4
```

Overall, basic calculations in R are similar to those in other programming
languages and follow the standard order of operations.

### Vectors

R has a variety of data types, including numbers, characters, and logical values
(`TRUE` or `FALSE`). For example, we can create a vector of numbers using the
`c()` function:

```{r}
x <- c(1, 2, 3, 4, 5)
```

We can also create a vector of characters using the same function:

```{r}
y <- c("apple", "banana", "cherry")
```

R also has functions for performing mathematical operations on data. For
example, we can use the `mean()` function to calculate the average of a vector
of numbers, or we can use the `length()` function to calculate the number of
elements in a vector:

```{r}
mean(x)
length(y)
```

### Matrices

Matrices are two-dimensional arrays of data, and can be created using the
`matrix()` function. For example, the following code creates a `3x3` matrix of
random numbers:

```{r}
mat <- matrix(rnorm(9), nrow = 3, ncol = 3)
```

### Data Frames

Data frames are similar to matrices, but can contain different types of data in
each column. Data frames can be created using the `data.frame()` function. For
example, the following code creates a data frame containing three types of
data: 

```{r}
df <- data.frame(x = c(1, 2, 3), y = c("a", "b", "c"), z = c(TRUE, FALSE, TRUE))
```

In addition to these basic features, R has many packages and libraries that
extend its capabilities and provide additional functions and tools for data
analysis and visualization. Some popular packages include `dplyr` for data
manipulation, `ggplot2` for data visualization, and `caret` for machine
learning. With its versatility and robust community, R is a valuable tool for
anyone working with data.

## Packages

A package is a small bundle of code that an R user wrote and uploaded to a
central server for anybody to access, download and use.

### Installing a Package

In many cases, you will download an R package from the
[CRAN](http://cran.r-project.org/) server. This can be done via the
`install.packages()` function by providing the package name in quotes. For
example, an important package for data visualization is `ggplot2`. To install
this package, you can simply run:

```{r, eval=FALSE}
install.packages("ggplot2")
```

This book illustrates its concepts and ideas using the **jfa** package, an R
package for statistical auditing, which can also be downloaded from CRAN. Before
running the examples in this book, you should install this package by running
the following command in R:

```{r, eval=FALSE}
install.packages("jfa")
```

##$ Loading a Package

Once you have installed a package, you must load it into every R session. To
load a package into your R session, call `library()` and provide the name of the
package (without quotes) that you want to load. For example, before running the
examples in this book, you can load the **jfa** package with:

```{r, eval=FALSE}
library(jfa)
```

> If you want a deeper understanding of how the **jfa** package works, or want
to look at the source code of the package, see the
[package website](https://koenderks.github.io/jfa/).

### Updating a Package

R packages are updated regularly. To update a package in your R library you
should call `update.packages()` and provide the name of the package that you
want to update in quotes. For example, each time there is a new release of the
**jfa* package, you can update it by running:

```{r, eval=FALSE}
update.packages("jfa")
```

## Loading Data

### Loading Data from a CSV File

A commonly used data type is a `.csv` file. You can load this type of files
via the `read.csv()` function. For example, if the file `example.csv` is in
the current working directory, you can load it by running:

```{r, eval=FALSE}
read.csv("example.csv")
```

### Loading Data from an Excel File

Another commonly used data type are Excel files. You can load this type of files
via the `read_excel()` function from the `readxl` package. For this to work, you
should first install this package using the `install.packages()` command and
load it into the R session using a call to `library()`. For example, if the file
`example.xlsx` is in the current working directory, and the data you want to
load is on the first worksheet, you can load it by running:

```{r, eval=FALSE}\
install.packages("readxl")
library(readxl)
read_excel("example.csv", sheet = 1)
```
