---
output:
  bookdown::html_document2:
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r echo = FALSE, cache = FALSE}
source("utils.R", local = TRUE)
```

# R Basics {#CHAPTER-R-BASICS}

This chapter covers the basics: installing and using packages and loading data.

## Installing a Package

This book illustrates its concepts and ideas using the `jfa` package, which can
be downloaded from CRAN. You can install this package by running the command
below in R:

```{r, eval=FALSE}
install.packages("jfa")
```


## Loading a Package

To load a package into your R session, call `library()` and provide the name of
the package you want to load. Naturally, this requires that you already have the
package installed on the computer. For example, before running the examples in
this book, you can load the `jfa` package with:

```{r, eval=FALSE}
library(jfa)
```

> If you want a deeper understanding of how the `jfa` package works, or want to
look at the source code of the package, see the
[package website](https://koenderks.github.io/jfa/).

Packages in R are collections of functions and/or data that are bundled up for easy distribution, and installing a package will extend the functionality of R on your computer. If an R user creates a package and thinks that it might be useful for others, that user can distribute it through a package repository. The primary repository for distributing R packages is called CRAN (the Comprehensive R Archive Network), but there are others, such as Bioconductor, which specializes in packages related to genomic data.

If you have spent much time learning R, you may have heard of the *tidyverse*, which is a collection of R packages that share common ideas of how data should be structured and manipulated. This is in contrast to *base R*, which is the set of packages that are included when you just download and install R. The tidyverse is a set of add-ons for R, which make it easier to do many operations related to data manipulation and visualization. This book mostly uses the tidyverse, as I believe that it provides a quicker and simpler (but not less powerful!) way to work with data.

## Upgrading a Package

To update a package in your R library, call `update.packages()` and provide the
name of the package you want to update. For example, you can update the `jfa`
package with:

```{r, eval=FALSE}
update.packages("jfa")
```


## Loading a Delimited Text Data File


## Loading Data from an Excel File
