---
output:
  bookdown::html_document2:
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r echo = FALSE, cache = FALSE}
source("utils.R", local = TRUE)
```

# R Basics {#CHAPTER-R-BASICS}

This chapter covers the basics: installing and using packages and loading data.

## Installing a Package

A package is a small bundle of code that an R user wrote and uploaded to a
central server for anybody to access, download and use. In many cases, you will
download an R package from the [CRAN](http://cran.r-project.org/) server. This
can be done via the `install.packages()` function by providing the package name
in quotes. For example, an important package for data visualization is
`ggplot2`. To install this package, you can simply run:

```{r, eval=FALSE}
install.packages("ggplot2")
```

This book illustrates its concepts and ideas using the `jfa` package, which can
also be downloaded from CRAN. You can install this package by running the
following command in R:

```{r, eval=FALSE}
install.packages("jfa")
```

## Loading a Package

Once you have installed a package, you must load it into every R session. To
load a package into your R session, call `library()` and provide the name of the
package (without quotes) that you want to load. For example, before running the
examples in this book, you can load the `jfa` package with:

```{r, eval=FALSE}
library(jfa)
```

> If you want a deeper understanding of how the `jfa` package works, or want to
look at the source code of the package, see the
[package website](https://koenderks.github.io/jfa/).

## Upgrading a Package

R packages are updated regularly. To update a package in your R library you
should call `update.packages()` and provide the name of the package that you
want to update in quotes. For example, each time there is a new release of the
`jfa` package, you can update it by running:

```{r, eval=FALSE}
update.packages("jfa")
```

## Loading Data from a CSV File

A commonly used data type is a `.csv` file. You can load this type of files
via the `read.csv()` function. For example, if the file `example.csv` is in
the current working directory, you can load it by running:

```{r, eval=FALSE}
read.csv("example.csv")
```

## Loading Data from an Excel File

Another commonly used data type are Excel files. You can load this type of files
via the `read_excel()` function from the `readxl` package. For this to work, you
should first install this package using the `install.packages()` command and
load it into the R session using a call to `library()`. For example, if the file
`example.xlsx` is in the current working directory, and the data you want to
load is on the first worksheet, you can load it by running:

```{r, eval=FALSE}\
install.packages("readxl")
library(readxl)
read_excel("example.csv", sheet = 1)
```
