---
output:
  bookdown::html_document2:
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r echo = FALSE, cache = FALSE}
source("utils.R", local = TRUE)
```

# Evaluating a Stratified Sample {#CHAPTER-6}

In an audit context, stratified sampling can be used to select a sample of
transactions from different departments, locations, or business units to ensure
that the sample is representative of the population.

For example, if you want to audit the expense claims of a large organization,
you can stratify the population based on the department, location, or business
unit to ensure that all departments are represented accordingly in the sample.
Another example of such a situation would be a group audit where the audited
organization consists of different components or branches. Stratification is
relevant for the group auditor if they must form an opinion on the group as a
whole because they must aggregate the samples taken by the component auditors. 

As a data example, consider the `retailer` data set that comes with the package.
The organization in question consists of 20 branches across the country. In each
of the 20 strata, a component auditor has taken a statistical sample and
reported the outcomes to the group auditor. 

```{r}
data(retailer)
head(retailer)
```

In general, there are three approaches to evaluating a stratified sample: no
pooling, complete pooling, and partial pooling (see Derks et al., 2022). When
using `evaluation()`, you must to indicate which type of pooling to use via the
`pooling` argument. No pooling assumes no similarities between strata, which
means that all strata are analyzed independently. Complete pooling assumes no
difference between strata, which means that all data is aggregated and analyzed
as a whole. Finally, partial pooling assumes differences and similarities
between strata, which means that information can be shared between strata.
Partial pooling (i.e., multilevel/hierarchical modeling) is a powerful
technique that can result in more efficient population and stratum estimates but
is currently only feasible when performing a Bayesian analysis. For this reason,
this vignette only describes the Bayesian approach to stratified evaluation but
going from this approach to a classical approach only requires setting
`prior = FALSE`.

The number of units per stratum in the population can be provided with `N.units`
to weigh the stratum estimates to determine population estimate. This is called
poststratification. If `N.units` is not specified, each stratum is assumed to be
equally represented in the population.

## No pooling

No pooling (`pooling = "none"`, default) assumes no similarities between strata.
This means that the prior distribution specified through `prior` is applied
independently for each stratum. This allows for independent estimates for the
misstatement in each stratum but also results in a relatively high uncertainty
in the population estimate. The call below evaluates the sample using a Bayesian
stratified evaluation procedure, in which the stratum estimates are
poststratified to arrive at the population estimate.

```{r}
set.seed(1) # Important because the posterior distribution is determined via sampling
result_np <- evaluation(
  materiality = 0.05, method = "binomial", prior = TRUE,
  n = retailer$samples, x = retailer$errors, N.units = retailer$items,
  alternative = "two.sided", pooling = "none"
)
summary(result_np)
```

In this case, the output of the `summary()` function shows that the estimate of
the misstatement in the population is 5.85%, with the 95% credible interval
ranging from 4.28% to 8.22%. The stratum estimates differ substantially from
each other but are relatively uncertain.

```{r}
plot(result_np, type = "estimates")
```

The prior and posterior distribution for the population misstatement can be
requested via the `plot()` function.

```{r}
plot(result_np, type = "posterior")
```

## Complete pooling

Complete pooling (`pooling = "complete"`) assumes no differences between strata.
This has the advantages that data from all strata can be aggregated, which
decreases the uncertainty in the population estimate compared to the no
pooling approach. However, the disadvantage of this approach is that it does
not facilitate the distinction between between strata, as every stratum
receives the same estimate equal to that of the population. The call below
evaluates the sample using a Bayesian stratified evaluation procedure, in
which the strata are assumed to be the same.

```{r}
result_cp <- evaluation(
  materiality = 0.05, method = "binomial", prior = TRUE,
  n = retailer$samples, x = retailer$errors, N.units = retailer$items,
  alternative = "two.sided", pooling = "complete"
)
summary(result_cp)
```

For example, the output of the `summary()` function shows that the estimate of
the misstatement in the population is 4.47%, with the 95% credible interval
ranging from 3.74% to 5.33%. Since the data is aggregated, the stratum estimates
contain relatively little uncertainty. However, the probability of misstatement
in stratum 20 (many misstatements) under this assumption is the same as that of
stratum 15 (few misstatements).

```{r}
plot(result_cp, type = "estimates")
```

The prior and posterior distribution for the population misstatement can be
requested via the `plot()` function.

```{r}
plot(result_cp, type = "posterior")
```

## Partial pooling

Finally, partial pooling (`pooling = "partial"`) assumes differences and
similarities between strata. This allows the auditor to differentiate between
strata, while also sharing information between the strata to reduce uncertainty
in the population estimate. The call below evaluates the sample using a Bayesian
stratified evaluation procedure, in which the stratum estimates are
poststratified to arrive at the population estimate.

```{r}
set.seed(1) # Important because the posterior distribution is determined via sampling
result_pp <- evaluation(
  materiality = 0.05, method = "binomial", prior = TRUE,
  n = retailer$samples, x = retailer$errors, N.units = retailer$items,
  alternative = "two.sided", pooling = "partial"
)
summary(result_pp)
```

In this case, the output shows that the estimate of the misstatement in the
population is 4.34%, with the 95% credible interval ranging from 3.45% to 5.33%.
Note that this population estimate is substantially less uncertain than that of
the no pooling approach. Note that, like in the no pooling approach, the stratum
estimates are different from each other but lie closer together and are less
uncertain.

```{r}
plot(result_pp, type = "estimates")
```

The prior and posterior distribution for the population misstatement can be
requested via the `plot()` function.

```{r}
plot(result_pp, type = "posterior")
```

## Evaluation using data

For this example, we take the `allowances` that set that comes with the package.
This data set contains 3500 financial statement line items, each with a booked
value `bookValue` and, for illustrative purposes, and audited (true) value
`auditValue`. Since the focus of this vignette is the evaluation stage in the
audit, the sample is already indicated in the data set. The performance
materiality in this example is set to 5%.

```{r}
data(allowances)
head(allowances)
```

Evaluating a stratified sample using data requires specification of the `data`,
`values`, `values.audit` and `strata` arguments in the `evaluation()` function.
In this case, the units are monetary and calculated by aggregating the book
values of the items in each stratum.

```{r}
N.units <- aggregate(allowances$bookValue, list(allowances$branch), sum)$x
```

### Classical Evaluation

The call below evaluates the `allowances` sample using a classical stratified
evaluation procedure, in which the stratum estimates are poststratified to
arrive at the population estimate.

```{r}
x <- evaluation(
  materiality = 0.05, data = allowances,
  values = "bookValue", values.audit = "auditValue", strata = "branch", times = "times",
  alternative = "two.sided", N.units = N.units
)
summary(x)
```

In this case, the output shows that the estimate of the misstatement in the
population is 14.72%, with the 95% confidence interval ranging from 12.55% to
18.26%. The precision of the population estimate is 3.54%. The stratum estimates
can be seen in the output of the `summary()` function and are visualized below.

```{r}
plot(x, type = "estimates")
```

### Bayesian Evaluation

Bayesian inference can improve upon the estimates of the classical approach by
pooling information between strata where possible. The call below evaluates the
`allowances` sample using a Bayesian multilevel stratified evaluation procedure,
in which the stratum estimates are poststratified to arrive at the population
estimate.

```{r}
# x <- evaluation(
#   materiality = 0.05, data = allowances, prior = TRUE,
#   values = "bookValue", values.audit = "auditValue", strata = "branch", times = "times",
#   alternative = "two.sided", N.units = N.units, pooling = "partial"
# )
summary(x)
```

The output shows that the estimate of the misstatement in the population is
15.66%, with the 95% credible interval ranging from 14.59% to 17%. The precision
of the population estimate is 1.34%, which is substantially lower than that of
the classical approach. The stratum estimates can be seen in the output of the
`summary()` function and are visualized below.

```{r}
plot(x, type = "estimates")
```

The prior and posterior distribution for the population misstatement can be
requested via the `plot()` function.

```{r}
# plot(x, type = "posterior")
```
