```{r echo = FALSE, cache = FALSE}
source("utils.R", local = TRUE)
```

::: {.callout-note appearance="simple"}
You are reading the work-in-progress first edition of Statistical Auditing with
R. This chapter is currently a dumping ground of ideas, and it is incomplete.
:::

# R Packages {#chap-packages}

This chapter discusses other R packages implementing statistical techniques for
audit sampling.

## MUS

**MUS** [@MUS] is an R package providing sampling and evaluation methods to
apply Monetary Unit Sampling during an audit of financial statements. The
package is available via [CRAN](https://cran.r-project.org/package=MUS) and can
be downloaded by running the following code:

```{r, eval=FALSE}
install.packages("MUS")
```

Note that the **MUS** package provides no functionality for Bayesian audit
sampling.

## audit

**audit** [@audit] is an R package based on the paper by @Meeden2007 that can be
used to find an upper bound for the total amount of overstatement of assets in a
set of accounts and estimating the amount of sales tax owed on a collection of
transactions. The package is available via [CRAN](https://cran.r-project.org/package=audit)
and can be downloaded by running the following code:

```{r, eval=FALSE}
install.packages("audit")
```

Note that the **audit** package provides limited functionality for evaluating
audit samples and has no functionality for planning and selecting audit samples.

## samplingbook

**samplingbook** [@samplingbook] is an R package based on the book
**Stichproben: Methoden und praktische Umsetzung mit R** [@Kauermann2010] that
focuses on survey sampling and statistical analysis of these samples. The
package is available via [CRAN](https://cran.r-project.org/package=samplingbook)
and can be downloaded by running the following code:

```{r, eval=FALSE}
install.packages("samplingbook")
```

Note that the **samplingbook** package provides limited functionality for
auditing as it mainly focuses on survey sampling.
